---
title: "Data Summary and Extinction Calculations"
author: "Chris Terry"
output: 
  html_document: 
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document calculates the summary statistics used in the paper and generates the plots describing extinction points. 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(lme4)
library(knitr)

Obs<- read_csv('../InitialDataStorage/TubeDoc_CC3_ForDataEntry.csv') %>%
  filter( !is.na(PAL_FEMALE))

Temps <- read_csv('../InitialDataStorage/TubeDoc_CC3_ALL.csv') %>%
  select(Gen,  LINE_ID, TUBE_ID, IncubTemp )

TempTable = Temps %>%
  pivot_wider(id_cols = LINE_ID, names_from = Gen, values_from = IncubTemp)

PrevTempTable = Temps %>%
  mutate( PrevTemp = IncubTemp,
          GenNum = parse_number(Gen)+1) %>%
  select(-IncubTemp, -TUBE_ID, -Gen  )

LaggedFemales_PAL <- read_csv('../ProcessedData/LaggedFemales_PAL.csv')
LaggedFemales_PAN <- read_csv('../ProcessedData/LaggedFemales_PAN.csv')

```


```{r}
### Adding in initial condition
Gen1 = filter(Temps, Gen == 'G01')%>%
  mutate( PAL_FEMALE=3,  ### as added on 6 June
          PAL_MALE=2,
          PAN_FEMALE=0,
          PAN_MALE=0 ,
          NOTES = 'Initialisation')

bind_rows(Gen1, Obs) %>%
  separate(LINE_ID, into = c('Comp', 'Var', 'Rep'), remove = FALSE) %>%
  mutate( Var_Bin = str_sub(Var, 1,1),
          GenNum = parse_number(Gen)) %>%
  left_join(PrevTempTable, 
            by = c('GenNum', 'LINE_ID' ))-> Obs2

## Confirming that the competition treatments are not different at the G2 census (when they should be the same as competitors ony just added)
Obs2 %>%
  filter( Gen == 'G02') %>%
  lm( data = ., PAL_FEMALE ~ Comp) %>%
  summary

```

## Incubator Temperature Plots

```{r message=FALSE}

expand.grid( day = (0:11)/12, 
             GenNum = 1:10) -> dayintervals

Obs2 %>%
  filter(Var_Bin == 'V') %>%
  left_join(dayintervals, relationship = "many-to-many") %>%
  mutate( GenNum_int = GenNum + day) %>%
  ggplot( aes( x = GenNum_int, y = IncubTemp))+
  geom_line()+ 
  geom_point(aes(col = IncubTemp))+
  facet_wrap(~Var)+
  theme_classic()+
  scale_x_continuous(breaks= 1:10)+
  scale_color_viridis_c(option = 'A', name = 'Temperature', limits = c(22.5, 30))+
  guides(col='none')+
  theme(axis.text.x = element_text(size = 7),
        strip.background = element_blank())+
  ylab('')+
  xlab('Generation') -> VariableIncubTemps

Obs2 %>%
  filter(Var_Bin == 'S') %>%
  left_join(dayintervals, relationship = "many-to-many") %>%
  mutate( GenNum_int = GenNum + day) %>%
  ggplot( aes( x = GenNum_int, y = IncubTemp))+
  geom_line()+
  geom_point(aes(col = IncubTemp), size =5)+
  #  facet_wrap(~Var)+
  theme_classic()+
  scale_x_continuous(breaks= 1:10)+
  scale_color_viridis_c(option = 'A', name = 'Temperature', limits = c(22.5, 30))+
  ylab('Incubator Temperature')+
  xlab('Generation') +
  theme(legend.position = 'bottom') -> SteadyIncubTemps

plot_grid(SteadyIncubTemps, VariableIncubTemps,
          labels = c('Steady Rise', 'Variable Rise'),
          label_fontface = 'plain',
          scale =0.9, rel_widths = c(1,2) )

ggsave('../Figures/Fig1b_tempplots.png', width = 8, height = 5, bg = 'white')
```

# Summary Totals

```{R}
Obs2 %>%
  summarise( sum(PAL_MALE), sum(PAL_FEMALE),
             sum(PAN_MALE, na.rm = TRUE), 
             sum(PAN_FEMALE, na.rm = TRUE))

Obs2 %>%
  summarise(sum(PAL_FEMALE>0),
            sum(PAN_FEMALE>0, na.rm = TRUE))
```

## Average of Non-Zero Counts

```{r}
LaggedFemales_PAL%>%
  filter( PAL_FEMALE>0) %>%
  summarise(mean(PAL_FEMALE),
            sd(PAL_FEMALE) )

```


# Sex Ratio

```{r}

## Calculating expected interval
Interval95 = data.frame(SampleSize =1:100,
                        N_female = (1:100)/2,
                        Lower = qbinom(prob = 0.5, size =1:100, p = 0.025),
                        Upper = qbinom(prob = 0.5, size =1:100, p = 0.975))

Obs2 %>%
  filter(Gen !='G01') %>%
  filter( !(PAL_FEMALE==0 & PAL_MALE==0)) %>%
  ggplot(aes(PAL_FEMALE ,PAL_MALE) )+
  geom_point(alpha = 0.1)+
  geom_abline(slope = 1, intercept = 0)+
  geom_smooth()+
  geom_line(data = Interval95, aes( x = N_female, y = Lower), col = 2)+
  geom_line(data = Interval95, aes( x = N_female, y = Upper), col = 2)+
  theme_classic()+
  theme(plot.title = element_text(face = 'italic'))+
  xlab('Female Count')+
  ylab('Male Count')+
  ggtitle('D. pallidifrons')+
  coord_fixed(xlim = c(0,80),ylim = c(0,80),ratio = 1)-> PAL_SR

Obs2 %>%
  filter( !(PAN_FEMALE==0 & PAN_MALE==0)) %>%
  ggplot(aes(PAN_FEMALE ,PAN_MALE) )+
  geom_point(alpha = 0.1)+
  geom_abline(slope = 1, intercept = 0)+
  geom_smooth()+
  geom_line(data = Interval95, aes( x = N_female, y = Lower), col = 2)+
  geom_line(data = Interval95, aes( x = N_female, y = Upper), col = 2)+
  theme_classic() +
  theme(plot.title = element_text(face = 'italic'))+
  xlab('Female Count')+
  ylab('Male Count')+
  ggtitle('D. pandora')+
  coord_fixed(xlim = c(0,80),ylim = c(0,80),ratio = 1) -> PAN_SR

plot_grid(PAL_SR,PAN_SR )
ggsave('../Figures/SI2_SexRatio.png', width = 6, height =4)

```

# Population Trajectories

Just including females in these plots. 

```{r}
## PAL
Obs2 %>% 
  mutate(Competition = ifelse( Comp =='C', 'With Competition', 'Monoculture')) %>%
  mutate(Temperature = ifelse( Var_Bin =='V', 'Variable Rise', 'Steady Rise')) %>%
  ggplot( aes( x = GenNum, y = PAL_FEMALE,
               col = IncubTemp   , group = LINE_ID))+
  geom_line()+
  scale_x_continuous(breaks= 1:10)+
  scale_color_viridis_c(option = 'A', name = 'Temperature', limits = c(22.5, 30))+
  facet_grid(Competition~Temperature)+
  theme_classic()+
  theme(plot.title = element_text(face = 'italic'))+
  xlab('Generation')+
  ylab('Female count')+
  ggtitle('D. pallidifrons') -> FemalePAL_TS

# PAN
Obs2 %>% 
  mutate(Competition = ifelse( Comp =='C', 'With Competition', 'Monoculture')) %>%
  mutate(Temperature = ifelse( Var_Bin =='V', 'Variable Rise', 'Steady Rise')) %>%
  filter( Comp == 'C') %>%
  ggplot( aes( x = GenNum, y = PAN_FEMALE,
               col = IncubTemp   , group = LINE_ID))+
  geom_line()+
  scale_x_continuous(breaks= 1:10)+
  scale_color_viridis_c(option = 'A', name = 'Temperature', limits = c(22.5, 30))+
  facet_wrap(~Temperature)+
  xlab('Generation')+
  ylab('Female count')+
  theme_classic()+
  theme(plot.title = element_text(face = 'italic'))+
  theme(legend.position = 'bottom', legend.direction = 'horizontal')+
  ggtitle('D. pandora') -> FemalePAN_TS

### Combining
CombinedPlot <- plot_grid(FemalePAL_TS+guides(col='none'),
                          FemalePAN_TS,
                          ncol = 1, rel_heights = c(2,1.5))

ggsave(plot = CombinedPlot,'../Figures/SI1_TimeSeries_all_fem.png', height = 8, width=6)

```

# Extinction Point Distributions

NB using both males and females here for last seen data

```{r}
Obs2 %>%
  mutate( PAL_TOT= PAL_FEMALE + PAL_MALE, 
          PAL_Alive = PAL_TOT>0) %>%
  group_by(LINE_ID,Comp,Var,Var_Bin) %>%
  summarise(N_gens = sum(PAL_Alive)) %>%
  ungroup() %>%
  mutate(Competition = ifelse( Comp =='C', 'With Competition', 'Monoculture')) %>%
  mutate(Temperature = ifelse( Var_Bin =='V', 'Variable Rise', 'Steady Rise')) %>%
  mutate(Treatment =paste0(Comp, Var_Bin))-> LastSeenData

LastSeenData  %>%
  group_by(Competition,Temperature) %>%
  summarise(Mean_Gens = mean(N_gens),
            SD_Gens = sd(N_gens),
             Mean_Extinct = Mean_Gens+1,
            Lower66_ext = quantile(N_gens,1/6, na.rm=TRUE)+1,
            Upper66_ext = quantile(N_gens,5/6, na.rm=TRUE)+1, 
            SE = SD_Gens / sqrt(60))%>%
  mutate(xmin = Mean_Gens-SD_Gens,
         xmax = Mean_Gens+SD_Gens) -> MeanGens

MeanGens %>% kable()


LastSeenData%>%  
  ggplot()+
  scale_x_continuous(breaks= 2:10)+
  geom_histogram(aes( x = N_gens+1 ), binwidth = 1, fill = 'skyblue')+
  theme_classic()+
  theme(strip.background = element_blank(),
        strip.text = element_text(size = 10))+
  facet_wrap(~Temperature+Competition, nrow = 1)+
  geom_point(data = MeanGens, aes( x = Mean_Gens+1), y = 1)+
  geom_errorbarh(data = MeanGens,  aes( xmin = xmin+1, xmax = xmax+1, y = 1))+
  geom_text(data = MeanGens, 
            aes( x = Mean_Gens+1, label = signif(Mean_Gens+1,3)),
            y = 3, size = 8)+
  xlab('Generation Recorded Extirpated')+
  ylab('Count')

ggsave('../Figures/Fig1C_ExtirpHists.png', height = 4, width = 8)
```

## Statistical tests of Extiprations

```{r}
### Overall of impact of treatment
aov( N_gens~Treatment, data =LastSeenData) %>% summary

## Follow-up test of pairwise comparisons
TukeyHSD(aov( N_gens~Treatment, data =LastSeenData)) %>% plot

## Overall distinction between 'MS' and others
LastSeenData %>%
  mutate( IS_MS= Treatment=='MS') %>%
  lm( N_gens~IS_MS, data =.) %>% 
  summary
```

## Effect of trajectory of variability

```{r}
Mod_RE <- lmer(data = filter(LastSeenData, Var_Bin =='V'),
               N_gens~Treatment + (1|Var))

Mod_wo <- lm(data = filter(LastSeenData, Var_Bin =='V'),
             N_gens~Treatment)

anova( Mod_RE, Mod_wo) ## exact pathway seems to be somewhat influential 
summary(Mod_RE)

filter(LastSeenData, Var_Bin =='V') %>%
  group_by(Var) %>%
  summarise(MeanGens= mean(N_gens)) -> GensPerTempRoute

Temps %>%
  separate(LINE_ID, into = c('Comp', 'Var', 'Rep'), remove = FALSE) %>%
  select(Gen, Var, IncubTemp) %>%
  distinct() -> TempRoutes

### Possible Features of temperature routes

# Initial Value
# Mean of first 3, 
# Mean of 4-6
# Mean of 7-10

TempRoutes %>%
  mutate( GenNum = parse_number(Gen))%>%
  filter( Var != 'S') %>%
  group_by(Var) -> TR2

### Comparing the 12 routes, initial value is not a significant predictor of mean gen
GensPerTempRoute %>%
  left_join(summarise(filter(TR2,Gen   == 'G01' ),  InitialValue = IncubTemp)) %>%
  left_join(summarise(filter(TR2,GenNum <4),  Mean1 = mean(IncubTemp))) %>%
  left_join(summarise(filter(TR2,GenNum %in% c(4,5,6)),  Mean2 = mean(IncubTemp))) %>%
  left_join(summarise(filter(TR2,GenNum %in% c(7,8,9,10)),  Mean3 = mean(IncubTemp))) %>%
  select(-1) %>%
  lm(data=.,    MeanGens~InitialValue) %>%
  summary()

### Looking at the individual cases:
filter(LastSeenData, Var_Bin =='V') %>%
  left_join(filter(TempRoutes, Gen   == 'G01') ) %>%
  lm(data=.,    N_gens ~IncubTemp*Treatment ) %>%
  summary() 

## Across the full 120 trials, gen 1 temp does come out as significant (N_gens ~IncubTemp), but this fades when treatment also included. 

GensPerTempRoute %>% 
  mutate(Rank = rank(MeanGens)) %>%
  left_join(TempRoutes, by = 'Var') %>%
  mutate( GenNum = parse_number(Gen)) %>%
  ggplot(aes( x = GenNum, y = IncubTemp, 
              col = MeanGens, group=Var))+
  geom_line()+
  facet_wrap(~Rank)

```

## Extinctions plotted on trajectories

```{R}

filter(LastSeenData, Var_Bin =='V') %>%
  separate(LINE_ID, into = c('Comp', 'Var', 'Rep'), remove = FALSE) %>%
  group_by(Var,N_gens ) %>%
  mutate(loc =22+  (1:n())/3  )-> ExtinctionsToPlot
  
GensPerTempRoute %>% 
  left_join(TempRoutes, by = 'Var') %>%
  mutate( GenNum = parse_number(Gen)) %>%
  ggplot()+
  geom_line(aes( x = GenNum, y = IncubTemp, 
               group=Var))+
  facet_wrap(~Var) +
  geom_point(aes( x = N_gens +1, y = loc, col = Competition       ),
             data =ExtinctionsToPlot , shape = 4)+
  theme_classic()+
  ylab('Incubator Temperature')+
  xlab('Generation')+
  scale_color_discrete(name = 'Extinctions\nunder treatment:')

ggsave( '../Figures/SI3_VarTrajs.png', width = 7, height = 5)

```

# Session Info

```{r}
sessionInfo()

```

