---
title: "Model Fitting"
author: "Chris Terry"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document makes a first go at fitting a suite of different competition models to the gen-gen transition data. 

# Data Ingestion

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(brms)

source('../Scripts/Make_brms_script_PAL_FEM.R')
source('../Scripts/Make_brms_script_PAN_FEM.R')

LaggedFemales_PAL <- read_csv('../ProcessedData/LaggedFemales_PAL.csv')
LaggedFemales_PAN <- read_csv('../ProcessedData/LaggedFemales_PAN.csv')

```


`Make_brms_script_PAL_FEM.R` and `Make_brms_script_PAN_FEM.R` are both functions to write small R scripts that can build brms / STAN models in a modular way. 
This works by building each segment of the model required and saving the R script  in `AutoModels/Scripts/`.  

# Overall Model Framework

$$N_{i,t+1} = N_{i,t}   \times B(T_t) \times C(N_{i,t},N_{j,t})$$

## Thermal performance curve options

### Gaussian

$$ B(T) = B_{pk} \times exp [ -0.5\times \dfrac{|T-T_{pk}|}{a}^2   ] $$

### Taylor Sexton

$$ B(T) = B_{pk} \times \dfrac{-(T-T_{min})^4  + 2 \times(T-T_{min})^2 \times (T_{pk}-T_{min})^2 }{(T_{pk}-T_{min})^4}  $$
### Atkin 

$$ B(T) = B_0\times (a-b\times T)^{T/10}$$


### Simplified Briere 1

$$ B(T) = a \times (T-T_{min}) \times)  \sqrt{T_{max} -T} $$



### Simplified $\beta$ type

$$ B(T) = \rho \times (a - \frac{T}{10})\times (\frac{T}{10})^b $$

### Linear 

$$ B(T) =  a + b\times T $$

### Gaussian-Gompertz

$$B(T) = a \times exp\left[-exp[b(T-T_{pk}) -\theta ] - c(T-T_{pk})^2\right]$$


## Competition model 

Saturating multi-species models for modifications to the maximum per-capita reproductive rate due to competition - only one equilibrium.

### Beverton-Holt


$$C(N_{i,t},N_{j,t}) = \dfrac{1}{ 1+ \alpha_{ii} N_{i,t} + \alpha_{ij}N_{j,t} }$$


### Beverton-Holt logged


$$C(N_{i,t},N_{j,t}) = \dfrac{1}{ 1+ \alpha_{ii} log(N_{i,t}+1) + \alpha_{ij}log(N_{j,t}+1) }$$


### Lotka-Volterra (linear)


$$C(N_{i,t},N_{j,t}) =  1+ \alpha_{ii} N_{i,t} + \alpha_{ij}N_{j,t} $$

### Exponential Beverton holt

Model 6 Law and Watkinson 1987 https://www.jstor.org/stable/2260211 

$$C(N_{i,t},N_{j,t}) =  \dfrac{1}{ (1+ N_{i,t} + \alpha_{ij}N_{j,t} )^\beta}$$

# Female Only Fits

```{r}
Option_df <- expand_grid( TP = c('Briere'),                                                            
                          Comp =   c('BH', 'BHBeta', 'BHTEMP', 'BHTEMP1',
                                     'BHTEMP1Beta','BHTEMPBeta', 'BHTEMP1Log','BHTEMP1LogBeta'),
                          Err = c('Pois', 'NB', 'ZiNB', 'ZiNB_temp','ZiNB_shapetemp','ZiNB_shapecomp'),  
                          ModelName =NA)

Option_df2 <- expand_grid( TP = c('Briere'),   ### Just testing out this alternative                                                         
                          Comp =   c('BH'),
                          Err = c('Norm'),  
                          ModelName =NA)
```

##  PAL
```{R}
Option_df_PAL <- Option_df %>% 
  mutate(   Name = paste(paste0('TP.', TP)   ,
                         paste0('Comp.', Comp)   ,
                         paste0('Err.', Err)   ,
                         sep = '_') )
```

```{R eval = FALSE}
LaggedFemales_PAL <- filter(LaggedFemales_PAL, PrevTemp < 28) ## cutting out top degree of observations

## Make all the scripts
for(i in 1:nrow(Option_df) ){
  try({
    file <-Make_brms_script_PAL_FEM(i ,Option_df,  cores = 4, iter = 4000)
    Option_df_PAL$ModelName[i] <- file
  })
}
## Run the scripts
for(i in 1:nrow(Option_df) ){
  print(i)
  try(source(Option_df_PAL$ModelName[i]))
}
```

```{R}
map_df(1:nrow(Option_df), function(i, Option_df){
  # print(i)
  xxx  <- load( file= paste0('../AutoModels/Fits/FEMS_PAL_', Option_df_PAL$Name[i]))
  get(xxx)$criteria$loo$estimates %>%
    as.data.frame()%>%
    rownames_to_column('IC') %>%
    mutate(Name = Option_df_PAL$Name[i]) %>%
    return()
}, Option_df) -> LOO_PAL

map_dbl(1:nrow(Option_df), function(i, Option_df){
  #  print(i)
  xxx  <- load( file= paste0('../AutoModels/Fits/FEMS_PAL_', Option_df_PAL$Name[i]))
  get(xxx) %>%
    rhat() %>%
    max() %>%
    return()
}, Option_df) -> max_rhats_PAL


LOO_PAL %>%
  pivot_wider(id_cols = Name, names_from = IC, values_from = c(Estimate, SE)) %>%
  mutate( Max_Estimate_elpd_loo = max(Estimate_elpd_loo ),
          Relative_elpd_loo = Estimate_elpd_loo -Max_Estimate_elpd_loo) %>%
  left_join(  data.frame(Name = Option_df_PAL$Name , 
                         Max_rhat = max_rhats_PAL) ) %>%
  arrange( desc(Estimate_elpd_loo)) %>%
  separate(Name, sep = '_', into = c('Thermal', 'Competition', 'Error', 'ZiNBModifier'),remove = FALSE) -> PAL_COMPAR_TABLE


PAL_COMPAR_TABLE%>%
  select(Thermal:ZiNBModifier , Relative_elpd_loo, SE_elpd_loo , Max_rhat  ) %>%
  as.data.frame()


PAL_COMPAR_TABLE %>%
  select(Thermal:ZiNBModifier , Relative_elpd_loo, Max_rhat  ) %>%
  mutate( Competition = str_sub(Competition, 6, -1), 
          Thermal  = str_sub(Thermal, 4, -1), 
          Error  = str_sub(Error, 5, -1), 
          Relative_elpd_loo = signif(Relative_elpd_loo,3), 
          Max_rhat = signif(Max_rhat,4))  %>%
  write_csv('../Tables/PAL_LooCompare.csv')
```

### Comparing best model

```{R}
# load( file='../AutoModels/Fits/FEMS_PAL_TP.Briere_Comp.BHEXPTEMPBeta_Err.ZiNB_temp')
# PAL_TP.Briere_Comp.BHEXPTEMPBeta_Err.ZiNB_temp_fit
# PAL_TP.Briere_Comp.BHEXPTEMPBeta_Err.ZiNB_temp_fit %>% plot

load( file='../AutoModels/Fits/FEMS_PAL_TP.Briere_Comp.BHTEMP1Beta_Err.ZiNB_temp')
load( file='../AutoModels/Fits/FEMS_PAL_TP.Briere_Comp.BHTEMP1_Err.ZiNB_temp')

loo_compare(PAL_TP.Briere_Comp.BHTEMP1Beta_Err.ZiNB_temp_fit, 
            PAL_TP.Briere_Comp.BHTEMP1_Err.ZiNB_temp_fit)

PAL_TP.Briere_Comp.BHTEMP1_Err.ZiNB_temp_fit

```

## Fitting PAN 

```{R}
Option_df_PAN <- Option_df %>% 
  mutate(   Name = paste(paste0('TP.', TP)   ,
                         paste0('Comp.', Comp)   ,
                         paste0('Err.', Err)   ,
                         sep = '_') )
```

### Running Pandora full fit

```{R eval = FALSE}

## Make all the scripts
for(i in 1:nrow(Option_df) ){
  try({
    file <-Make_brms_script_PAN_FEM(i ,Option_df,  cores = 4, iter = 4000)
    Option_df_PAN$ModelName[i] <- file
  })
}
for(i in 1:nrow(Option_df) ){
  print(i)
  try(source(Option_df_PAN$ModelName[i]))
}
```

```{R}

map_df(1:nrow(Option_df), function(i, Option_df){
  # print(i)
  xxx  <- load( file= paste0('../AutoModels/Fits/FEMS_PAN_', Option_df_PAN$Name[i]))
  get(xxx)$criteria$loo$estimates %>%
    as.data.frame()%>%
    rownames_to_column('IC') %>%
    mutate(Name = Option_df_PAN$Name[i]) %>%
    return()
}, Option_df) -> LOO_PAN


map_dbl(1:nrow(Option_df), function(i, Option_df){
  #  print(i)
  xxx  <- load( file= paste0('../AutoModels/Fits/FEMS_PAN_', Option_df_PAN$Name[i]))
  get(xxx) %>%
    rhat() %>%
    max() %>%
    return()
}, Option_df) -> max_rhats_PAN


LOO_PAN %>%
  pivot_wider(id_cols = Name, names_from = IC, values_from = c(Estimate, SE)) %>%
  mutate( Max_Estimate_elpd_loo = max(Estimate_elpd_loo ),
          Relative_elpd_loo = Estimate_elpd_loo -Max_Estimate_elpd_loo) %>%
  left_join(  data.frame(Name = Option_df_PAN$Name , 
                         Max_rhat = max_rhats_PAN) ) %>%
  arrange( desc(Estimate_elpd_loo)) %>%
  separate(Name, sep = '_', into = c('Thermal', 'Competition', 'Error', 'ZiNBModifier'),remove = FALSE) -> PAN_COMPAR_TABLE

PAN_COMPAR_TABLE%>%
  select(Thermal:ZiNBModifier , Relative_elpd_loo, SE_elpd_loo , Max_rhat  ) %>%
  as.data.frame



PAN_COMPAR_TABLE %>%
  select(Thermal:ZiNBModifier , Relative_elpd_loo, Max_rhat  ) %>%
  mutate( Competition = str_sub(Competition, 6, -1), 
          Thermal  = str_sub(Thermal, 4, -1), 
          Error  = str_sub(Error, 5, -1), 
          Relative_elpd_loo = signif(Relative_elpd_loo,3), 
          Max_rhat = signif(Max_rhat,4))  %>%
  write_csv('../Tables/PAN_LooCompare.csv')
```

### Viewing best model

```{R}
#load( file='../AutoModels/Fits/FEMS_PAN_TP.Briere_Comp.BHEXPTEMP_Err.ZiNB_shapecomp')
#PAN_TP.Briere_Comp.BHEXPTEMP_Err.ZiNB_shapecomp_fit
#PAN_TP.Briere_Comp.BHEXPTEMP_Err.ZiNB_shapecomp_fit %>% plot

```

# Session Info

```{r}
sessionInfo()

```

