---
title: "Predictions From Direct Simulation Of Models"
author: "Chris Terry"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script takes the best-fitting models, and runs stochastic simulations from them to estimate the distribution of extinction points. While it assumes a particular model form, there is no particular theoretical framework. 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(brms)
library(knitr)
library(cowplot)

## Best Models
load('../AutoModels/KeyFits/FEMS_PAL_TP.Briere_Comp.BHTEMP1_Err.ZiNB_temp')
load('../AutoModels/KeyFits/FEMS_PAN_TP.Briere_Comp.BHTEMP1_Err.ZiNB_shapecomp')

## Poisson Models
load( file = '../AutoModels/Fits/FEMS_PAN_TP.Briere_Comp.BHTEMP1_Err.Pois')
load( file=  '../AutoModels/Fits/FEMS_PAL_TP.Briere_Comp.BHTEMP1_Err.Pois')

```

# Main Models

The `brms` models will output a full posterior of predictions each time. 1000 draws are used in parallel. To traverse the timesteps, the same parameter draw is used for each step. 

-Draws are taken using `posterior_predict`, which incorporates noise from the error distribution.
-Each PAL population starts equally. 
-Pandora's are added as per experimental regime. 

## Preamble
```{R}
N_rep= 1000
N_Gens = 20

## Pallidifrons initiation
PAL_df <- as.data.frame(matrix(NA, nrow = N_rep, ncol = N_Gens ))
colnames(PAL_df) <- paste0('Gen_', 1:N_Gens)
PAL_df$ID = 1:N_rep
PAL_df$Gen_1 <- 3

## Pandora initiation
PAN_df <- as.data.frame(matrix(NA, nrow = N_rep, ncol = N_Gens ))
colnames(PAN_df) <- paste0('Gen_', 1:N_Gens)
PAN_df$ID = 1:N_rep
PAN_df$Gen_1 <- 0

## Temperature 
Steady_rise = matrix(seq(24, by= 0.4, l = N_Gens), nrow=1)

VarRise <- matrix(Steady_rise, nrow = N_rep,ncol = N_Gens, byrow = TRUE) +
  matrix(sample(c(0, +1.5, - 1.5),
                replace = TRUE, size = N_rep*N_Gens),
         nrow = N_rep,ncol = N_Gens)

set.seed(1)
BRMS_DRAWS <- sample(1:8000, size = N_rep, replace = FALSE) ## To mix across chains
```


## Simulation Function

```{r}

Simulator<-function(Temps, PAN_df, PAL_df, N_rep,N_Gens , 
                    COMP=TRUE,
                    Model = c('Best', 'Poisson')[1] ){
  
  set.seed(1)
  
  Scenario <- paste( ifelse( nrow(Temps)==1,
                             'Steady',
                             'Var' ), 
                     COMP,Model,
                     sep = '_')
  
  for(G in 1:(N_Gens-1)){
    
    InputData                      <- data.frame( ID = 1:N_rep)  ## build up data.frame for putting into models
    InputData$Prev_PAL_FEMALE      <- PAL_df[,G]
    InputData$Prev_PAN_FEMALE      <- PAN_df[,G]
    InputData$PrevTemp             <- Temps[,G]
    InputData$Temp20               <- InputData$PrevTemp -20
    
    ## Pandora Addition
    if(COMP & G == 2){InputData$Prev_PAN_FEMALE =  InputData$Prev_PAN_FEMALE + 8}
    if(COMP &  G == 4){InputData$Prev_PAN_FEMALE =  InputData$Prev_PAN_FEMALE + 4}
    if(COMP &  G == 8){InputData$Prev_PAN_FEMALE =  InputData$Prev_PAN_FEMALE + 4}
    
    ## Calculate numbers for competition modelling
    InputData <- mutate(InputData, 
                        Prev_PAN_FEMALE_comp = if_else(Prev_PAN_FEMALE==0,
                                                       0,Prev_PAN_FEMALE-1 ))
    InputData <- mutate(InputData, 
                        Prev_PAL_FEMALE_comp = if_else(Prev_PAL_FEMALE==0,
                                                       0,Prev_PAL_FEMALE-1 ))
    
    if( Model == 'Best'){
      ### output from posterior crosses inputs and draws - need 
      
      PAL_df[,(G+1)] <-   diag(posterior_predict(PAL_TP.Briere_Comp.BHTEMP1_Err.ZiNB_temp_fit, 
                                           newdata = InputData,draw_ids = BRMS_DRAWS))
      PAN_df[,(G+1)] <-  diag(posterior_predict(PAN_TP.Briere_Comp.BHTEMP1_Err.ZiNB_shapecomp_fit, 
                                           newdata = InputData, draw_ids = BRMS_DRAWS))  
    }
    if(Model == 'Poisson'){
      PAL_df[,(G+1)] <-  diag(posterior_predict(PAL_TP.Briere_Comp.BHTEMP1_Err.Pois_fit, 
                                           newdata = InputData, draw_ids = BRMS_DRAWS))  
      
      PAN_df[,(G+1)] <-  diag(posterior_predict(PAN_TP.Briere_Comp.BHTEMP1_Err.Pois_fit, 
                                           newdata = InputData, draw_ids = BRMS_DRAWS))
    }
  }
  
  ## Plotting ( for debugging)
  PLOT <- PAL_df %>%
    pivot_longer(cols = -ID,names_to= 'Gen', values_to = 'NumPAL',
                 names_prefix = 'Gen_', names_transform = as.integer) %>%
    ggplot(aes( x = Gen, y = NumPAL, col=factor(ID)))+
    geom_point()+ geom_line()+guides( col = 'none')

  
  PAL_df %>%
    pivot_longer(cols = -ID,names_to= 'Gen', values_to = 'NumPAL',
                 names_prefix = 'Gen_', names_transform = as.integer) %>%
    mutate(NumPAL = ifelse(is.nan(NumPAL), 0, NumPAL),
           PAL_Alive = NumPAL>0) %>%
    group_by(ID) %>%
    summarise(N_gens = sum(PAL_Alive, na.rm=TRUE)) %>%
    mutate(Scenario =  Scenario) %>%
    ungroup() -> Num_Gens_Alive
  ### ^^NB Not quite the same as the extinction data used as here just modelling females, but very close 
  return(list('PAL_df'=PAL_df, 'PAN_df'= PAN_df ,'PLOT'= PLOT, 
              'Num_Gens_Alive'=Num_Gens_Alive))  
}

```

# Running

## Steady Rise +Comp
```{r eval = FALSE}
SR_out <- Simulator(Temps= Steady_rise,
                    PAN_df, PAL_df, N_rep,N_Gens , COMP = TRUE)

SR_out$PLOT
```

## Var Rise +Comp
```{r eval = FALSE}
VR_out <- Simulator(Temps= VarRise,
                    PAN_df, PAL_df, N_rep,N_Gens , COMP = TRUE)

VR_out$PLOT
```

## No Competition

```{R eval = FALSE}
SR_nocomp_out <- Simulator(Temps= Steady_rise,
                           PAN_df, PAL_df, N_rep,N_Gens, COMP = FALSE )



mean((SR_nocomp_out$Num_Gens_Alive$N_gens)+1)  

VR_nocomp_out <- Simulator(Temps= VarRise,
                           PAN_df, PAL_df, N_rep,N_Gens, COMP = FALSE )
```

## Level of zero-inflation with temperature

Direct simulation methods dominated by details of temperature dependence of zero-inflation. 

```{r}
temp <- seq(22, 28, by = 0.1)

probZERO_mid  = inv_logit_scaled( -17.43  + temp *  0.63   )
plot( probZERO_mid ~temp)

DRAWS<-as_draws_df(PAL_TP.Briere_Comp.BHTEMP1_Err.ZiNB_temp_fit)
probZERO_100  = inv_logit_scaled( DRAWS$b_zi_Intercept[1:100]  + rep(temp, 100) *  DRAWS$b_zi_PrevTemp[1:100])
plot( probZERO_100 ~rep(temp, 100))

```

# Repeat with Poisson Model

## Steady Rise +Comp
```{r eval = FALSE}
SR_out_POIS <- Simulator(Temps= Steady_rise,
                         PAN_df, PAL_df, N_rep,N_Gens,COMP = TRUE,  Model = 'Poisson' )

SR_out_POIS$PLOT
SR_out_POIS$Num_Gens_Alive$N_gens %>% hist(main = 'Number of Generations seen')

## Generation Extinct

mean((SR_out_POIS$Num_Gens_Alive$N_gens)+1)  
quantile((SR_out_POIS$Num_Gens_Alive$N_gens)+1, probs = c(0.025, 1/3, 0.5, 2/3, 0.975))
```

## Var Rise +Comp
```{r eval = FALSE}
VR_out_POIS <- Simulator(Temps= VarRise,
                         PAN_df, PAL_df, N_rep,N_Gens ,COMP = TRUE,  Model = 'Poisson')

VR_out_POIS$PLOT

VR_out_POIS$Num_Gens_Alive$N_gens %>% hist(main = 'Number of Generations seen', breaks= 1:15)

```

## No Competition

```{R eval = FALSE}
SR_nocomp_out_POIS  <- Simulator(Temps= Steady_rise,
                           PAN_df, PAL_df, N_rep,N_Gens, COMP = FALSE ,  Model = 'Poisson' )

## Populations are higher without competition, but extinction point is pretty similar. 
SR_out$PAL_df %>% colMeans()
SR_nocomp_out$PAL_df %>% colMeans() 

VR_nocomp_out_POIS  <- Simulator(Temps= VarRise,
                           PAN_df, PAL_df, N_rep,N_Gens, COMP = FALSE ,  Model = 'Poisson')
```

# Saving

```{r eval = FALSE}
All_Simulation_list <- list('SR_out'= SR_out ,
                            'VR_out'= VR_out,
                            'SR_nocomp_out'= SR_nocomp_out,
                            'VR_nocomp_out'= VR_nocomp_out,
                            'VR_out_POIS'= VR_out_POIS ,
                            'SR_out_POIS'= SR_out_POIS,
                            'SR_nocomp_out_POIS'= SR_nocomp_out_POIS,
                            'VR_nocomp_out_POIS'= VR_nocomp_out_POIS)

save(All_Simulation_list, file = '../ProcessedData/All_Simulation_list')

All_Simulation_list %>%
  map_df('Num_Gens_Alive') %>%
  mutate(N_gen_extinct = N_gens+1) %>%
  write_csv( '../Tables/AllSimulationNumGensAlive.csv')
```

```{R}
load(file = '../ProcessedData/All_Simulation_list')
```

```{r}
VR_out <- All_Simulation_list$VR_out
VR_nocomp_out<- All_Simulation_list$VR_nocomp_out
## Populations are higher without competition
## with competition
VR_out$Num_Gens_Alive$Scenario[1]
VR_out$PAL_df %>% colMeans()

## without competition
VR_nocomp_out$Num_Gens_Alive$Scenario[1]
VR_nocomp_out$PAL_df %>% colMeans() 

#  extinction point is pretty similar, but in the 'wrong' direction!

list(VR_out, VR_nocomp_out) %>%
  map_df('Num_Gens_Alive')%>%
  group_by(Scenario ) %>%
  mutate(N_gen_extinct = N_gens+1) %>%
summarise(Mean=mean(N_gen_extinct))

plot_grid(VR_out$PLOT,
         VR_nocomp_out$PLOT, ncol = 1,
          labels = c( 'Var Rise, Competition, Best Model',
                      'Var Rise,  Monoculture, Best Model'))

full_join(VR_out$PAN_df  %>%
            pivot_longer(cols = -ID,names_to= 'Gen', values_to = 'NumPAN',
                         names_prefix = 'Gen_',
                         names_transform = as.integer) %>%
            mutate( PreviousPAN = lag(NumPAN)),
          VR_out$PAL_df  %>%
            pivot_longer(cols = -ID,names_to= 'Gen', values_to = 'NumPAL',
                         names_prefix = 'Gen_',
                         names_transform = as.integer) %>%
            mutate( PreviousPAL = lag(NumPAL)))%>%
 mutate(PreviousPAN = ifelse(is.na(PreviousPAN), 0, PreviousPAN)) -> FullDF_CompComparison


FullDF_CompComparison %>%
  #filter( ID<20) %>%
  ggplot( aes(col = PreviousPAL, y = NumPAL, x = 1/(1+PreviousPAN)  ))+
  geom_point()+
  geom_smooth( se = FALSE)+
  scale_x_log10()+
  ggtitle( 'Pandora competition defintely reduces growth|')

FullDF_CompComparison %>%
  #filter( ID<20) %>%
  ggplot( aes(y = NumPAL, x = 1/(1+PreviousPAL)  ))+
  geom_point()+
  geom_smooth( se = FALSE)+
  scale_x_log10()+
  ggtitle( '')

FullDF_CompComparison %>%
  left_join(All_Simulation_list[[1]]$Num_Gens_Alive) %>%
  filter( Gen == N_gens)
  

```


# Table

```{r}
All_Simulation_list %>%
  map_df('Num_Gens_Alive')%>%
  group_by(Scenario ) %>%
  mutate(N_gen_extinct = N_gens+1) %>%
  mutate( Temp_Extinct = (N_gen_extinct*0.4 + 23.6 )) %>%
  summarise(Mean=mean(N_gen_extinct), 
            LQ95= quantile(N_gen_extinct, 0.025 ),
            HQ95= quantile(N_gen_extinct, 0.975),
            LQ66= quantile(N_gen_extinct, 1/6 ),
            HQ66= quantile(N_gen_extinct, 5/6 )
  ) %>%
  as.data.frame()%>%
  separate(Scenario, into = c('Environment', 'Competition', 'Model')) %>%
  arrange( Model, Environment, desc(Competition))%>%
  kable()

```

# Session Info

```{R}
sessionInfo()
```

