---
title: "8 Refitting with less experiment reuse"
author: "Chris Terry"
date: '`r Sys.Date()`'
output: 
  html_document: 
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{R message=FALSE, warning=FALSE}
library(tidyverse)
library(brms)
library(posterior)
logit = boot::logit
inv.logit = boot::inv.logit
Temp_grid <- expand.grid(Temp = seq(22.5,28.7, by =0.1),  .draw = 1:1000)

```

The core models presented in the main body have an element of circularity - they are ultimately tested on the observed extinction of PAL observed in  the same experiments the generation-to-generation population dynamics were fit on. 

The data can be split into distinct test-and-train data with varying levels of strictness, with various implications for the amount of available data. 

Excluding PAL observations from the Steady/Competition treatment for the model fitting is the most direct approach, done here. This keeps ~75% of PAL growth data, but only 50% of data on the competitive pressure of PAN on PAL 

Note:

- there is still some information leakage because the original model selection used this data. 

- PAN is still fit using data from the eventual test experiment. 

```{r}
load('../AutoModels/KeyFits/FEMS_PAL_TP.Briere_Comp.BHTEMP1_Err.ZiNB_temp')
LaggedFemales_PAL <- read_csv('../ProcessedData/LaggedFemales_PAL.csv')
TRAIN_Data<-  filter(LaggedFemales_PAL,
                     str_sub( LINE_ID, 1, 3) != 'C_S') ## exclude transistions from test set. 

```

## Refitting model
```{R eval = FALSE}
PAL_TP.Briere_Comp.BHTEMP1_Err.ZiNB_temp_fit %>%
  update(newdata = TRAIN_Data) -> PAL_fit_without_test_data

save(PAL_fit_without_test_data,file =  '../AutoModels/KeyFits/PAL_fit_without_test_data' )
```

# PAN
```{r}
load( file = '../AutoModels/Fits/FEMS_PAN_TP.Briere_Comp.BHTEMP1_Err.ZiNB_shapecomp')
PAN_BEST_FIT <-    PAN_TP.Briere_Comp.BHTEMP1_Err.ZiNB_shapecomp_fit 
```

##PAL
```{r}
load(  '../AutoModels/KeyFits/PAL_fit_without_test_data')

PAL_fit_without_test_data %>%
  as_draws_df() %>%
  thin_draws(thin = 8)%>%
  as.data.frame() %>% 
  left_join(Temp_grid) %>%
  mutate( Temp20 = Temp-20, 
          Base_lambda= b_B0_Intercept *  Temp * (Temp - b_Tmin_Intercept) * sqrt(b_Tmax_Intercept - Temp) ,
          Zi_adjust = 1-inv.logit(b_zi_Intercept + Temp*b_zi_PrevTemp)  ,
          lambda = (Base_lambda *Zi_adjust)  )%>%
  mutate(PAL_alphaii = b_alphaii0_Intercept+ b_alphaiiT_Intercept*Temp20,
         PAL_alphaij = b_alphaij_Intercept )  %>%
  select(PAL_lambda = lambda,
         Temp,
         Zi_adjust,
         PAL_alphaii, 
         PAL_alphaij, 
         .draw) %>%
  mutate(PAL_K = (PAL_lambda-1)/PAL_alphaii) -> PAL_Params_NON_CIRCULAR

```

### Comparing Parameters in models

#### All-Data Approach

```{R}
PAL_TP.Briere_Comp.BHTEMP1_Err.ZiNB_temp_fit %>%
  as_draws_df() %>%
  thin_draws(thin = 8)%>%
  as.data.frame() %>% 
  left_join(Temp_grid) %>%
  mutate( Temp20 = Temp-20, 
          Base_lambda= b_B0_Intercept *  Temp * (Temp - b_Tmin_Intercept) * sqrt(b_Tmax_Intercept - Temp) ,
          Zi_adjust = 1-inv.logit(b_zi_Intercept + Temp*b_zi_PrevTemp)  ,
          lambda = (Base_lambda *Zi_adjust)  )%>%
  mutate(PAL_alphaii = b_alphaii0_Intercept+ b_alphaiiT_Intercept*Temp20,
         PAL_alphaij = b_alphaij_Intercept )  %>%
  select(PAL_lambda = lambda,
         Temp,
         Zi_adjust,
         PAL_alphaii, 
         PAL_alphaij, 
         .draw) %>%
  mutate(PAL_K = (PAL_lambda-1)/PAL_alphaii) -> PAL_Params_Alldata

```


```{R}
PAL_Params_NON_CIRCULAR$Data <- 'Independent'
PAL_Params_Alldata$Data <- 'All'

bind_rows(PAL_Params_NON_CIRCULAR, 
          PAL_Params_Alldata) %>%
  select(-PAL_K) %>%
  pivot_longer( cols = c(PAL_lambda, Zi_adjust, PAL_alphaii ,PAL_alphaij),
                names_to = 'Parameter', values_to = 'Value' ) %>%
  ggplot( aes( x =Temp, col = Parameter, y = Value))+
  geom_point( size = 0.1, alpha = 0.1)+
  geom_smooth()+
  facet_grid(Parameter~Data, scales = 'free_y')

```

##PAN

(same as main doc)

```{R}
PAN_BEST_FIT %>%
  as_draws_df() %>%
  thin_draws(thin = 8)%>%
  as.data.frame() %>%
  left_join(Temp_grid) %>%
  mutate(Temp20 = Temp-20, 
         Base_lambda =  b_B0_Intercept *  Temp * (Temp - b_Tmin_Intercept) * sqrt(b_Tmax_Intercept - Temp) ,
         Zi_adjust =  1-zi,
         lambda = (Base_lambda *Zi_adjust)   )%>%
  mutate(PAN_alphaii = b_alphaii_Intercept,
         PAN_alphaij = b_alphaij0_Intercept+ b_alphaijT_Intercept*Temp20)  %>%
  select(PAN_lambda = lambda,
         Temp, .draw,
         PAN_alphaii,PAN_alphaij)-> PAN_Params
```

# Calculating Coexistence across Temperatures


## Inferring parameters 

```{r}
alpha_FLOOR = 0.01
lambda_floor = 0.01

##  PAL_Params_NON_CIRCULAR$PAL_alphaij <- PAL_Params_NON_CIRCULAR$PAL_alphaij*0.5 ## to test impact of altering parameters

full_join(PAL_Params_NON_CIRCULAR, PAN_Params, 
          by = c('Temp', '.draw')) %>%
  mutate(PAL_alphaii = ifelse(PAL_alphaii>alpha_FLOOR, PAL_alphaii, alpha_FLOOR), 
         PAL_alphaij = ifelse(PAL_alphaij>alpha_FLOOR, PAL_alphaij, alpha_FLOOR),
         PAN_alphaii = ifelse(PAN_alphaii>alpha_FLOOR, PAN_alphaii, alpha_FLOOR), 
         PAN_alphaij = ifelse(PAN_alphaij>alpha_FLOOR, PAN_alphaij, alpha_FLOOR))  %>%
  mutate( PAL_lambda  = ifelse(PAL_lambda <lambda_floor | is.nan(PAL_lambda ),
                               lambda_floor, PAL_lambda   )) %>% 
  mutate( FitnessDiff =  ((PAN_lambda) / (PAL_lambda) )*sqrt( (PAL_alphaii*PAL_alphaij ) / (PAN_alphaii*PAN_alphaij)   ),
          NicheOverlap = sqrt( (PAN_alphaij*PAL_alphaij ) / (PAN_alphaii*PAL_alphaii)),
          NicheDiff = 1-NicheOverlap) -> CoexistQuants
```

```{R}

CoexistQuants %>%
  filter( Temp <=28, Temp >= 24) %>%
  mutate(PAN_Wins =  FitnessDiff > 1/NicheOverlap    &   FitnessDiff > NicheOverlap  ,
         PAL_Wins = FitnessDiff < NicheOverlap       &   FitnessDiff < 1/NicheOverlap , 
         Coexist = !PAN_Wins & !PAL_Wins   &  NicheDiff>0 , 
         PE  = !PAN_Wins & !PAL_Wins   &  NicheDiff<0 ) %>%          #  filter( Temp==25.5)%>%
  select( Temp, PAN_Wins:PE) %>%
  group_by(Temp) %>%
  summarise( M_PAN= mean(PAN_Wins, na.rm = TRUE)*100,
             M_PAL=mean(PAL_Wins, na.rm = TRUE)*100,
             M_Cex=mean(Coexist, na.rm = TRUE)*100,
             M_1PE = mean(PE, na.rm=TRUE)*100) %>%    
  pivot_longer(names_to = 'Outcome', values_to = 'Frac', cols = starts_with('M_')) %>% 
  ggplot(aes( x = Temp))+
  geom_bar(aes(y = Frac, fill = Outcome), stat = 'identity', width = 0.1)+
  scale_fill_manual(values = c('M_PAN' = 'lightpink2','M_PAL' = 'steelblue2',
                               'M_Cex' = 'grey50','M_1PE' = 'grey90'))+
  theme_classic()+
  ylab('Percentage of posterior\nwith each outcome')+
    geom_vline(xintercept = 27.4, col = 'gold', linewidth =3)+
  xlab('Temperature')+
  scale_x_continuous(sec.axis = sec_axis(~  ./0.4 - 59 , breaks = c(1:10),
                                         name = 'Generation'))+
  geom_label(x = 26.1, y = 80, label = 'Coexistence' ,fontface = 'italic',
             label.r=unit(0, 'lines'),   label.padding = unit(0.75, "lines"))+
  geom_label(x = 27.25, y = 25, label = 'Exclusion of\nD. pallidifrons'  ,fontface = 'italic',
             label.r=unit(0, 'lines'),  label.padding = unit(0.75, "lines"))+
  geom_label(x = 24.6, y = 40, label = 'Exclusion of\nD. pandora'  ,fontface = 'italic',
             label.r=unit(0, 'lines'),  label.padding = unit(0.75, "lines"))+
  guides(fill = 'none')-> Predictions_Plot

ggsave( filename = '../Figures/SI7_NonCircularPredictions.png', 
          height = 5, width = 6)

Predictions_Plot
```

## Temperature Prediction Range

```{r}
CoexistQuants %>%
  filter( Temp <=29, Temp >= 24) %>%
  mutate(PAN_Wins =  FitnessDiff > 1/NicheOverlap    &   FitnessDiff > NicheOverlap  ,
         PAL_Wins = FitnessDiff < NicheOverlap       &   FitnessDiff < 1/NicheOverlap , 
         Coexist = !PAN_Wins & !PAL_Wins   &  NicheDiff>0 , 
         PE  = !PAN_Wins & !PAL_Wins   &  NicheDiff<0 ) %>%          #  filter( Temp==25.5)%>%
  select( Temp, PAN_Wins:PE) %>%
  group_by(Temp) %>%
  summarise( M_PAN= mean(PAN_Wins, na.rm = TRUE)*100) %>%    
  pivot_longer(names_to = 'Outcome', values_to = 'Frac', cols = starts_with('M_')) -> ExclusionDF

## Searching for median and 66% central interval  (0.1666   - 0.83333)
## Inferring to nearest half

## Linear Interpolation 

 ## interpolate temperature: 
interp<- approx( y = ExclusionDF$Temp, x = ExclusionDF$Frac, xout = 100*c( 1/6, 0.5, 5/6)  )
interp


```

# Session Info

```{r}
sessionInfo()

```








